Classical Planning with Simulators: \\ Results on the Atari Video Games

### 2015-04-30 ijcai15-paper/paper.tex
Mods for Camera Ready
* Added UniMelb cluster data
* Added data on the UCT experiments
* Added authors and affiliations

\author{\textbf{ID:} 320  \ \ ; \ \ \textbf{Keywords:} classical planning, search, video games, application}

% \author{Nir Lipovetzky \\
%        The University of Melbourne \\
%     Melbourne, Australia\\
%     {\normalsize\url{nir.lipovetzky@unimelb.edu.au}}
%%% footnote doesn't look nice ; this looks better
%% although can't use Nir's unimelb address that is too long :-)
%% 
%    {\small\url{nirlipo@gmail.com}}
%    \And
%    Miquel Ramirez \\
%    Australian National University \\
%    Canberra, Australia\\
%   {\small\url{miquel.ramirez@gmail.com}}
%    \And
%    Hector Geffner \\
%    ICREA \&  Universitat Pompeu Fabra \\
%    Barcelona, SPAIN \\
%    {\small\url{hector.geffner@upf.edu}}
 %%%%% \thanks{firstname.lastname}
%}

@@

We tested \IW(1) and \BFS\ over \numgames\  of the $55$ different games 
considered in \cite{bellemare:jair2013}, from now on  abbreviated as BNVB.\footnote{
We left out  \textsc{Skiing} as the reported figures apparently use a  different reward structure.}
%% check
% \footnote{
% Entries marked as  ``--'' in the table are   still running at the time of submission. 
% Each run can take up to $24$ hours of CPU time and there are $10$ runs per game and algorithm. 
% Eventually all of  the  $55$ games will  be included in the evaluation.}
The two  algorithms were used to play the  games in the \emph{on-line planning
setting} supported by ALE where we will compare them with the planning algorithms 
considered by BNVB; namely, breadth-first search and \UCT. ALE supports also a \emph{learning setting} 
where the goal is to learn controllers that map  states into actions without
doing any lookahead. Algorithms across  the two settings are thus not directly comparable
as they compute different things. Learning controllers appears as a more challenging problem
and it is thus not surprising that planning algorithms like \UCT\ tend to achieve a higher score than 
learning algorithms. In addition, the learning algorithms reported by BNVB tend to use 
the state of the screen pixels, while the planning algorithms,  use the state
of the RAM memory. It is not clear however whether one input representation is
more challenging than the other; for the learning algorithms, BNVB mention
that the results tend to be better for the screen inputs.

@@

Last, in the lookahead, \IW(1) and \BFS\ select an action every $5$ frames, while \UCT\  selects 
an action in every frame.  This means that in order to explore a branch 
$300$ frames deep, UCT gets to choose $300$ actions, while \IW($1$) and \BFS\ 
get to choose $60$ actions, both however using the same $300$ frames from the
budget. For this, we  followed the setup  of \BRFS\ in BNVB 
that also selects actions every $5$ frames, matching the behavior
of the emulator that requests an action also every $5$ frames. Since the 
lookahead budget is given by a maximum number of (new) frames, and the time 
is mostly taken by calls to the emulator, this may not be the best choice  
for $\IW(1)$ and \BFS\ that may therefore not be exploiting all the 
options afforded by the budget. We'll look at this further in the future.

@@

\newpage

>>>>

%\author{\textbf{ID:} 320  \ \ ; \ \ \textbf{Keywords:} classical planning, search, video games, application}

 \author{Nir Lipovetzky \\
        {\small The University of Melbourne }\\
     {\small Melbourne, Australia}\\
     {\footnotesize\url{nir.lipovetzky@unimelb.edu.au}}
%% footnote doesn't look nice ; this looks better
% although can't use Nir's unimelb address that is too long :-)
% 
%    {\tiny\url{nirlipo@gmail.com}}
    \And
    Miquel Ramirez \\
    {\small Australian National University} \\
    {\small Canberra, Australia}\\
   {\footnotesize\url{miquel.ramirez@anu.edu.au}}
    \And
    Hector Geffner \\
    {\small ICREA \& Universitat Pompeu Fabra }\\
    {\small Barcelona, Spain }\\
    {\footnotesize\url{hector.geffner@upf.edu}}
 %%%% \thanks{firstname.lastname}
}

@@

We tested \IW(1) and \BFS\ over \numgames\  of the $55$ different games 
considered in \cite{bellemare:jair2013}, from now on  abbreviated as BNVB.\footnote{
We left out  \textsc{Skiing} as the reported figures apparently use a  different reward structure.}
%% check
% \footnote{
% Entries marked as  ``--'' in the table are   still running at the time of submission. 
% Each run can take up to $24$ hours of CPU time and there are $10$ runs per game and algorithm. 
% Eventually all of  the  $55$ games will  be included in the evaluation.}
The two  algorithms were used to play the  games in the \emph{on-line planning
setting} supported by ALE where we will compare them with the planning algorithms 
considered by BNVB; namely, breadth-first search and \UCT. ALE supports also a \emph{learning setting} 
where the goal is to learn controllers that map  states into actions without
doing any lookahead. Algorithms across  the two settings are thus not directly comparable
as they compute different things. Learning controllers appears as a more challenging problem
and it is thus not surprising that planning algorithms like \UCT\ tend to achieve a higher score than 
learning algorithms. In addition, the learning algorithms reported by BNVB tend to use 
the state of the screen pixels, while the planning algorithms,  use the state
of the RAM memory. It is not clear however whether one input representation is
more challenging than the other; for the learning algorithms, BNVB mention
that the results tend to be better for the screen inputs. Experiments were
run on a high-performance computation cluster run by the University of Melbourne, where each computing node consists of a $6$-core Intel Xeon 
E5-2440, with $2.4$ GHz clock speed, with 64 GBytes of RAM installed.

@@

Last, in the lookahead, \IW(1) and \BFS\ select an action every $5$ frames, while \UCT\  selects 
an action in every frame.  This means that in order to explore a branch 
$300$ frames deep, UCT gets to choose $300$ actions, while \IW($1$) and \BFS\ 
get to choose $60$ actions, both however using the same $300$ frames from the
budget. For this, we  followed the setup  of \BRFS\ in BNVB 
that also selects actions every $5$ frames, matching the behavior
of the emulator that requests an action also every $5$ frames. Since the 
lookahead budget is given by a maximum number of (new) frames, and the time 
is mostly taken by calls to the emulator, this may not be the best choice  
for $\IW(1)$ and \BFS\ that may therefore not be exploiting all the 
options afforded by the budget. Interestingly, when \UCT\ is limited to one
action each $5$ frames, its performance is reduced by up to a $50$\% in
games where it performs very well (\CrazyClimber), and does not a difference in those games where it does very poorly (\Freeway). We'll look at this further in the future.

@@
